---
title: "MiLB game logs data"
author: "Ronit Gandhi"
date: "2024-06-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CODE TO GET DATA FOR EACH AND EVERY GAME (PER GAME PITCH-BY-PITCH DATA) and SQL insert query (TEST)
```{r}
## CODE TO GET DATA FOR EACH AND EVERY GAME (PER GAME PITCH-BY-PITCH DATA) and SQL insert query (TEST)
library(baseballr)
library(dplyr)
library(DBI)
library(RMySQL)
library(data.table)

#DB connection
#con <- dbConnect(RMySQL::MySQL(), dbname = 'baseball_data', host = 'localhost', port = 3306, user = 'root', password = 'Gandhi_op16', local_infile = 1)
#message("connection success!")

# Sample game from 2023 season
payload <- get_pbp_mlb(723719)  

# Converting the tibble into a dataframe
names(payload) <- gsub("\\.", "_", names(payload))

payload_df <- as.data.frame(payload)

print(payload_df)
#checking the column names
colnames(payload_df)

#tryCatch({
 #   dbWriteTable(con, name = "game_details", value = payload_df, append = TRUE, row.names = FALSE)
#}, error = function(e) {
 #   print(e)
#})
write.csv(payload_df, "paylaod_df_base_review_inprog.csv", row.names = FALSE)


```
## CODE TO GET GAMES DATA FOR AAA for any season

```{r}
## CODE TO GET GAMES DATA FOR AAA for any season
library(baseballr)
library(dplyr)

# Function to get game data for a given date
get_game_data_for_date <- function(date_str, level_id = 11) {
  games <- tryCatch(
    {
      get_game_pks_mlb(date = date_str, level_ids = level_id)
    },
    error = function(e) {
      message("Error fetching data for ", date_str, ": ", e$message)
      return(NULL)
    }
  )
  return(games)
}

# Define the date range for the season
start_date <- as.Date("2019-03-01")
end_date <- as.Date("2019-10-30")
dates <- seq.Date(start_date, end_date, by = "day")

# Create a dataframe from the dates
dates_df <- data.frame(date = dates)

# Initialize an empty data frame to store all game data
all_game_data <- data.frame()

# Loop through each date in the dataframe and fetch game data
for (i in 1:nrow(dates_df)) {
  date_str <- format(dates_df$date[i], "%Y-%m-%d")
  message("Fetching data for ", date_str)
  daily_games <- get_game_data_for_date(date_str)
  
  # Check if daily_games is not NULL and is a data frame
  if (!is.null(daily_games) && is.data.frame(daily_games) && nrow(daily_games) > 0) {
    all_game_data <- bind_rows(all_game_data, daily_games)
  }
}

# Save the combined data to a CSV file
write.csv(all_game_data, "aaa_games_2019.csv", row.names = FALSE)
print("Game data extraction complete. CSV file saved as 'aaa_games_2019.csv'")

```

## CODE TO GET PBP DATA FOR ALLGAMES IN THE 2023 SEASON, AND INSERTED IN DATABASE
```{r}
## CODE TO GET PBP DATA FOR ALLGAMES IN THE 2023 SEASON, AND INSERTED IN DATABASE
library(baseballr)
library(dplyr)
library(DBI)
library(RMySQL)
library(data.table)

# DB connection
con <- dbConnect(RMySQL::MySQL(), dbname = 'baseball_data', host = 'localhost', port = 3306, user = 'root', password = 'Gandhi_op16', local_infile = 1)
message("connection success!")

# Load game IDs from CSV
game_ids <- fread("C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/Data with umpires/aaa_games_2023_with_umpires.csv")$game_pk

# Loop over game IDs and fetch and insert data
for (game_id in game_ids) {
    message(paste("Fetching data for game ID:", game_id))
    tryCatch({
        payload <- get_pbp_mlb(game_id)
        
        names(payload) <- gsub("\\.", "_", names(payload))
        if("reviewDetails_isOverturned" %in% names(payload)) {
            names(payload)[names(payload) == "reviewDetails_isOverturned"] <- "reviewDetails_isOverturned_x"
        }
        if("reviewDetails_inProgress" %in% names(payload)) {
            names(payload)[names(payload) == "reviewDetails_inProgress"] <- "reviewDetails_inProgress_x"
        }
        if("reviewDetails_challengeTeamId" %in% names(payload)) {
            names(payload)[names(payload) == "reviewDetails_challengeTeamId"] <- "reviewDetails_challengeTeamId_x"
        }
        if("reviewDetails_reviewType" %in% names(payload)) {
            names(payload)[names(payload) == "reviewDetails_reviewType"] <- "reviewDetails_reviewType_x"
          }
        payload_df <- as.data.frame(payload)
        
        # Check if the dataframe is not empty and has columns before attempting to write to the database
        if (ncol(payload_df) > 0 && nrow(payload_df) > 0) {
            dbWriteTable(con, name = "game_details", value = payload_df, append = TRUE, row.names = FALSE)
            print(paste("Data inserted for game ID:", game_id))
        } else {
            print(paste("No data fetched for game ID:", game_id))
        }
    }, error = function(e) {
        print(paste("Error fetching data for game ID:", game_id, "Error:", e$message))
    })
}
 
# Disconnect from DB
dbDisconnect(con)
message("Disconnected from database.")

```

## Basic EDA of Pith Data
pitch_data <- fread("C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data.csv", fill = TRUE, quote = "", check.names = FALSE)
```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)

# Step 1: Load the data
# Adjust the path to your CSV file
pitch_data <- fread('C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data.csv', fill = TRUE, quote = "", check.names = FALSE)

# Convert relevant columns to numeric, handling non-numeric values gracefully
pitch_data <- pitch_data %>%
  mutate(
    pitchData_startSpeed = as.numeric(pitchData_startSpeed),
    pitchData_endSpeed = as.numeric(pitchData_endSpeed),
    pitchData_zone = as.numeric(pitchData_zone),
    pitchData_typeConfidence = as.numeric(pitchData_typeConfidence),
    pitchData_plateTime = as.numeric(pitchData_plateTime),
    pitchData_extension = as.numeric(pitchData_extension),
    pitchData_coordinates_aY = as.numeric(pitchData_coordinates_aY),
    pitchData_coordinates_aZ = as.numeric(pitchData_coordinates_aZ),
    pitchData_coordinates_pfxX = as.numeric(pitchData_coordinates_pfxX),
    pitchData_coordinates_pfxZ = as.numeric(pitchData_coordinates_pfxZ),
    pitchData_coordinates_pX = as.numeric(pitchData_coordinates_pX),
    pitchData_coordinates_pZ = as.numeric(pitchData_coordinates_pZ),
    pitchData_coordinates_vX0 = as.numeric(pitchData_coordinates_vX0),
    pitchData_coordinates_vY0 = as.numeric(pitchData_coordinates_vY0),
    pitchData_coordinates_vZ0 = as.numeric(pitchData_coordinates_vZ0),
    pitchData_coordinates_x0 = as.numeric(pitchData_coordinates_x0),
    pitchData_coordinates_y0 = as.numeric(pitchData_coordinates_y0),
    pitchData_coordinates_z0 = as.numeric(pitchData_coordinates_z0),
    pitchData_coordinates_aX = as.numeric(pitchData_coordinates_aX),
    pitchData_breaks_breakAngle = as.numeric(pitchData_breaks_breakAngle),
    pitchData_breaks_breakLength = as.numeric(pitchData_breaks_breakLength),
    pitchData_breaks_breakY = as.numeric(pitchData_breaks_breakY),
    pitchData_breaks_spinRate = as.numeric(pitchData_breaks_spinRate),
    pitchData_breaks_spinDirection = as.numeric(pitchData_breaks_spinDirection)
  )

# Inspect the data to ensure conversion
print("Data summary after conversion:")
print(summary(pitch_data))
print(head(pitch_data))
# Get the unique game_pk values
game_ids <- unique(pitch_data$game_pk)

# Set up the plotting area to show multiple plots at a time
par(mfrow=c(2, 2))  # Adjust as needed

# Step 2: Iterate over each game and create visualizations
for (game_id in game_ids) {
  game_data <- pitch_data %>% filter(game_pk == game_id)
  

  print(game_data)
  # Plot: Scatter plot of Start Speed vs End Speed
  plot_scatter <- ggplot(game_data, aes(x = pitchData_startSpeed, y = pitchData_endSpeed)) +
    geom_point(alpha = 0.5) +
    theme_minimal() +
    labs(title = paste("Start Speed vs End Speed for Game", game_id), x = "Start Speed (mph)", y = "End Speed (mph)")
  
  print(plot_scatter)
  
  # Plot: Histogram of Start Speed
  plot_hist_start_speed <- ggplot(game_data, aes(x = pitchData_startSpeed)) +
    geom_histogram(bins = 30, fill = "blue", color = "black") +
    theme_minimal() +
    labs(title = paste("Distribution of Start Speed for Game", game_id), x = "Start Speed (mph)", y = "Frequency")
  
  print(plot_hist_start_speed)
  
  # Plot: Histogram of End Speed
  plot_hist_end_speed <- ggplot(game_data, aes(x = pitchData_endSpeed)) +
    geom_histogram(bins = 30, fill = "green", color = "black") +
    theme_minimal() +
    labs(title = paste("Distribution of End Speed for Game", game_id), x = "End Speed (mph)", y = "Frequency")
  
  print(plot_hist_end_speed)
  
  # Plot: Bar plot of Zone
  plot_bar_zone <- ggplot(game_data, aes(x = factor(pitchData_zone))) +
    geom_bar(fill = "purple", color = "black") +
    theme_minimal() +
    labs(title = paste("Distribution of Zone for Game", game_id), x = "Zone", y = "Count")
  
  print(plot_bar_zone)
  
  # Additional plots can be added here for other columns as needed
}

print("Plots have been displayed.")

```

## Batted Balls
```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)
library(baseballr)

# Step 1: Load the data
# Adjust the path to your CSV file
pitch_data <- fread('C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data.csv', fill = TRUE, quote = "", check.names = FALSE)

# Convert relevant columns to numeric, handling non-numeric values gracefully
pitch_data <- pitch_data %>%
  mutate(
    pitchData_coordinates_x = as.numeric(pitchData_coordinates_x),
    pitchData_coordinates_y = as.numeric(pitchData_coordinates_y),
    batted_ball_result = as.factor(batted_ball_result)  # Assuming batted_ball_result is a factor
  )

# Define the color palette
bb_palette <- c('Single' = "#006BA4",
                'Double' = "#A2CEEC", 
                'Triple'= "#FFBC79", 
                'Home Run'= "#C85200", 
                'Out/Other' = "#595959")

# Function to plot spray chart for each game_pk
plot_spray_chart <- function(game_id, game_data) {
  ggplot(game_data, aes(x = pitchData_coordinates_x, y = -pitchData_coordinates_y, color = batted_ball_result)) +
    geom_point(size = 3) +
    scale_color_manual(values = bb_palette) +
    labs(title = paste("Batted Balls: Game", game_id),
         x = "Coordinate X",
         y = "Coordinate Y",
         color = "Result") +
    theme_minimal()
}

# Get the unique game_pk values
game_ids <- unique(pitch_data$game_pk)

# Step 2: Iterate over each game and create visualizations
par(mfrow=c(2, 2))  # Adjust as needed
print(game_ids)
for (game_id in game_ids) {
  game_data <- pitch_data %>% filter(game_pk == game_id)
  
  # Plot the spray chart
  plot_scatter <- plot_spray_chart(game_id, game_data)
  print(plot_scatter)
}

print("Plots have been displayed.")

```

## Hit Data
```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)
library(baseballr)

# Step 1: Load the data
# Adjust the path to your CSV file
pitch_data <- fread('C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data.csv', fill = TRUE, quote = "", check.names = FALSE)

# Convert relevant columns to numeric and factor, handling non-numeric values gracefully
pitch_data <- pitch_data %>%
  mutate(
    hitData_coordinates_coordX = as.numeric(hitData_coordinates_coordX),
    hitData_coordinates_coordY = as.numeric(hitData_coordinates_coordY),
    hitData_trajectory = as.factor(hitData_trajectory),
    hitData_hardness = as.factor(hitData_hardness),
    hitData_location = as.factor(hitData_location)
  )

# Inspect the data to ensure conversion
print("Data summary after conversion:")
print(summary(pitch_data))

# Check the unique values in hitData_trajectory and hitData_hardness
unique_trajectories <- unique(pitch_data$hitData_trajectory)
unique_hardness <- unique(pitch_data$hitData_hardness)

print("Unique values in hitData_trajectory:")
print(unique_trajectories)

print("Unique values in hitData_hardness:")
print(unique_hardness)

# Define the color palette for trajectory based on the actual data
trajectory_colors <- c("#006BA4", "#A2CEEC", "#FFBC79", "#C85200", "#595959", "#7F7F7F", "#B6D7A8", "#76A5AF", "#C27BA0", "#FFD966")
trajectory_palette <- setNames(trajectory_colors[1:length(unique_trajectories)], unique_trajectories)

# Define the color palette for hardness based on the actual data
hardness_colors <- c("#006BA4", "#A2CEEC", "#FFBC79", "#C85200", "#595959", "#7F7F7F", "#B6D7A8", "#76A5AF", "#C27BA0", "#FFD966")
hardness_palette <- setNames(hardness_colors[1:length(unique_hardness)], unique_hardness)

# Function to plot spray chart for each game_pk
plot_spray_chart <- function(game_id, game_data) {
  ggplot(game_data, aes(x = hitData_coordinates_coordX, y = -hitData_coordinates_coordY, color = hitData_trajectory, shape = hitData_hardness)) +
    geom_point(size = 3) +
    scale_color_manual(values = trajectory_palette, na.translate = FALSE) +
    scale_shape_manual(values = 1:length(unique_hardness), na.translate = FALSE) +  # Different shapes for hardness levels
    labs(title = paste("Batted Balls: Game", game_id),
         x = "Coordinate X",
         y = "Coordinate Y",
         color = "Trajectory",
         shape = "Hardness") +
    theme_minimal()
}

# Get the unique game_pk values
game_ids <- unique(pitch_data$game_pk)

# Step 2: Iterate over each game and create visualizations
par(mfrow=c(2, 2))  # Adjust as needed

for (game_id in game_ids) {
  game_data <- pitch_data %>% filter(game_pk == game_id)
  
  # Remove rows with NA values
  game_data <- game_data %>% drop_na(hitData_coordinates_coordX, hitData_coordinates_coordY, hitData_trajectory, hitData_hardness)
  
  # Plot the spray chart
  plot_scatter <- plot_spray_chart(game_id, game_data)
  print(plot_scatter)
}

print("Plots have been displayed.")

```

## Robot Umpires Split
C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data.csv
```{r}
# Load necessary libraries
library(data.table)  # For fread and data manipulation

# Load the data with fill=TRUE to handle rows with varying field numbers
file_path <- 'C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data.csv'  # Adjust the path as necessary
pitch_data <- fread(file_path, fill = TRUE)

# Check for any obvious issues with the data right after loading
print("Data overview:")
str(pitch_data)

# Assuming there's a 'game_date' column, we need to convert it to the correct date format
# This step may need adjustment based on your actual date format
pitch_data[, game_date := as.Date(game_date, format = "%Y-%m-%d")]

# Extract the day of the week from the game_date
pitch_data[, day_of_week := weekdays(game_date)]

# Classify games by umpire type based on the day of the week
pitch_data[, umpire_type := ifelse(day_of_week %in% c("Tuesday", "Wednesday", "Thursday"),
                                 "Robot Umpire", "Human Umpire")]

# Output the first few rows to check the new column
print("Modified data preview:")
print(pitch_data)

# Optionally, save the modified data to a new CSV file
output_path <- 'C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data RobotSplit.csv'
fwrite(pitch_data, output_path)

print(paste("Modified data saved to", output_path))

```

## Do Pitchers pitch faster
```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)
library(ggplot2)

# Load the modified data
file_path <- 'C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data RobotSplit.csv' # Adjust the path as necessary
pitch_data <- fread(file_path, fill = TRUE)

# Convert pitch speed to numeric, ensuring no data issues interfere
pitch_data[, pitchData_startSpeed := as.numeric(pitchData_startSpeed)]

# Filter out any rows with NA in the relevant columns
pitch_data <- na.omit(pitch_data, cols = c("umpire_type", "pitchData_startSpeed"))

# Calculate summary statistics by umpire type
summary_stats <- pitch_data[, .(
  Average_Speed = mean(pitchData_startSpeed, na.rm = TRUE),
  Median_Speed = median(pitchData_startSpeed, na.rm = TRUE),
  SD_Speed = sd(pitchData_startSpeed, na.rm = TRUE)
), by = umpire_type]

print(summary_stats)

# Visualization: Boxplot of pitch speeds by umpire type
ggplot(pitch_data, aes(x = umpire_type, y = pitchData_startSpeed, fill = umpire_type)) +
  geom_boxplot() +
  labs(title = "Comparison of Pitch Speeds by Umpire Type",
       x = "Umpire Type",
       y = "Pitch Speed (mph)") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

# Visualization: Density Plot of Pitch Speeds by Umpire Type
ggplot(pitch_data, aes(x = pitchData_startSpeed, fill = umpire_type)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Pitch Speeds by Umpire Type",
       x = "Pitch Speed (mph)",
       y = "Density") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

```

## Do Pitchers spin the ball more

```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)
library(ggplot2)

# Load the modified data
file_path <- 'C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data RobotSplit.csv'  # Adjust the path as necessary
pitch_data <- fread(file_path, fill = TRUE)

# Ensure numeric conversion for analysis
pitch_data[, pitchData_breaks_spinRate := as.numeric(pitchData_breaks_spinRate)]
pitch_data[, pitchData_breaks_spinDirection := as.numeric(pitchData_breaks_spinDirection)]

# Remove any rows with NA in the relevant columns
pitch_data <- na.omit(pitch_data, cols = c("umpire_type", "pitchData_breaks_spinRate", "pitchData_breaks_spinDirection"))

# Calculate summary statistics by umpire type for spin rate
spin_rate_stats <- pitch_data[, .(
  Average_SpinRate = mean(pitchData_breaks_spinRate, na.rm = TRUE),
  SD_SpinRate = sd(pitchData_breaks_spinRate, na.rm = TRUE)
), by = umpire_type]

# Calculate summary statistics by umpire type for spin direction
spin_direction_stats <- pitch_data[, .(
  Average_SpinDirection = mean(pitchData_breaks_spinDirection, na.rm = TRUE),
  SD_SpinDirection = sd(pitchData_breaks_spinDirection, na.rm = TRUE)
), by = umpire_type]

# Output the statistics
print("Spin Rate Statistics by Umpire Type:")
print(spin_rate_stats)

print("Spin Direction Statistics by Umpire Type:")
print(spin_direction_stats)

# Visualization: Comparing Spin Rate by Umpire Type
ggplot(pitch_data, aes(x = umpire_type, y = pitchData_breaks_spinRate, fill = umpire_type)) +
  geom_boxplot() +
  labs(title = "Comparison of Pitch Spin Rates by Umpire Type",
       x = "Umpire Type",
       y = "Spin Rate (rpm)") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

# Visualization: Comparing Spin Direction by Umpire Type
ggplot(pitch_data, aes(x = umpire_type, y = pitchData_breaks_spinDirection, fill = umpire_type)) +
  geom_boxplot() +
  labs(title = "Comparison of Pitch Spin Directions by Umpire Type",
       x = "Umpire Type",
       y = "Spin Direction (degrees)") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

```

## Hitting data analysis, do they hit harder or where do they hit?
```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)
library(ggplot2)

# Load the modified data
file_path <- 'C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data RobotSplit.csv'  # Adjust the path as necessary
pitch_data <- fread(file_path, fill = TRUE)

# Ensure numeric conversion for analysis
pitch_data[, hitData_launchSpeed := as.numeric(hitData_launchSpeed)]
pitch_data[, hitData_launchAngle := as.numeric(hitData_launchAngle)]
pitch_data[, hitData_totalDistance := as.numeric(hitData_totalDistance)]
pitch_data[, hitData_coordinates_coordX := as.numeric(hitData_coordinates_coordX)]
pitch_data[, hitData_coordinates_coordY := as.numeric(hitData_coordinates_coordY)]

# Remove any rows with NA in the relevant columns
pitch_data <- na.omit(pitch_data, cols = c("umpire_type", "hitData_launchSpeed", "hitData_launchAngle", "hitData_totalDistance", "hitData_coordinates_coordX", "hitData_coordinates_coordY"))

# Calculate summary statistics by umpire type for hitting data
hitting_stats <- pitch_data[, .(
  Avg_LaunchSpeed = mean(hitData_launchSpeed, na.rm = TRUE),
  Avg_LaunchAngle = mean(hitData_launchAngle, na.rm = TRUE),
  Avg_TotalDistance = mean(hitData_totalDistance, na.rm = TRUE),
  Avg_CoordX = mean(hitData_coordinates_coordX, na.rm = TRUE),
  Avg_CoordY = mean(hitData_coordinates_coordY, na.rm = TRUE)
), by = umpire_type]

print("Hitting statistics by umpire type:")
print(hitting_stats)

# Visualization: Comparing Launch Speed by Umpire Type
ggplot(pitch_data, aes(x = umpire_type, y = hitData_launchSpeed, fill = umpire_type)) +
  geom_boxplot() +
  labs(title = "Comparison of Launch Speed by Umpire Type",
       x = "Umpire Type",
       y = "Launch Speed (mph)") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

# Visualization: Density Plot of Launch Angle by Umpire Type
ggplot(pitch_data, aes(x = hitData_launchAngle, fill = umpire_type)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Launch Angles by Umpire Type",
       x = "Launch Angle (degrees)",
       y = "Density") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

ggplot(pitch_data, aes(x = hitData_totalDistance, fill = umpire_type)) +
  geom_histogram(bins = 30, alpha = 0.6) +
  labs(title = "Histogram of Total Hit Distances by Umpire Type",
       x = "Total Distance (feet)", y = "Frequency") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

# Visualization 2: Scatter Plot of Launch Speed vs. Launch Angle
ggplot(pitch_data, aes(x = hitData_launchSpeed, y = hitData_launchAngle, color = umpire_type)) +
  geom_point(alpha = 0.6) +
  labs(title = "Scatter Plot of Launch Speed vs. Launch Angle by Umpire Type",
       x = "Launch Speed (mph)", y = "Launch Angle (degrees)") +
  scale_color_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

# Visualization 3: Hexbin Plot of Coordinates where Hits Landed
ggplot(pitch_data, aes(x = hitData_coordinates_coordX, y = hitData_coordinates_coordY, fill = umpire_type)) +
  geom_hex() +
  labs(title = "Hexbin Plot of Hit Landing Coordinates by Umpire Type",
       x = "Coordinate X", y = "Coordinate Y") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal()

# More plots can be added for other metrics like Total Distance

```

## Different Pitch Types
```{r}
# Load necessary libraries
library(tidyverse)
library(data.table)
library(ggplot2)

# Load the data
file_path <- 'C:/Users/ronit/Desktop/RUTGERS/Jerry Kim - Summer project/Robot Umpires/Data/2023 pbp data RobotSplit.csv'  # Adjust the path as necessary
pitch_data <- fread(file_path, fill = TRUE)

# Ensure proper data type conversion for categorical analysis
pitch_data[, details_type_description := as.factor(details_type_description)]

# Remove any rows with NA values in key columns for clean plotting
pitch_data <- na.omit(pitch_data, cols = c("umpire_type", "details_type_description"))

# Calculate frequency of pitch types by umpire type
pitch_type_counts <- pitch_data[, .N, by = .(umpire_type, details_type_description)]
pitch_type_counts[, percentage := N / sum(N), by = umpire_type]

# Visualization: Bar Plot of Pitch Types by Umpire Type
ggplot(pitch_type_counts, aes(x = details_type_description, y = percentage, fill = umpire_type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Distribution of Pitch Types by Umpire Type",
       x = "Pitch Type", y = "Percentage") +
  scale_fill_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Adjust text for better visibility

# Optional: Advanced visualization with speed or spin rate
# For this, ensure you have numeric conversions for speed/spin rate as shown in previous examples
ggplot(pitch_data, aes(x = details_type_description, y = pitchData_startSpeed, color = umpire_type)) +
  geom_boxplot() +
  labs(title = "Pitch Speed Distribution by Pitch Type and Umpire Type",
       x = "Pitch Type", y = "Pitch Speed (mph)") +
  scale_color_manual(values = c("Robot Umpire" = "blue", "Human Umpire" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```